Models of Lambek calculus with subexponentials

### 2019-06-24 PaperDraft.tex
updates in the draft

Also, we have a multiplicative conjunction or non-commutative tensor. This connective plays the role of concatenation.

@@

@@

After that, we introduce inference polymodal inference rules for non-commutative subexponentials. One can apply substructural
rules only if there is the relevant index on the current modality. Also, the right introduction rule is a sort of the
generalised right $\Box$-introduction a la $K4$. Modality $!_s$ might be introduced on the right only if its index is less than any other subexponential index from the antecedent. That is, if we have already used stronger modality, then we may
apply the weaker one.

@@

Let us consider the current proof-theoretical and algorithmic results on Lambek calculus with additives and subexponentials.
The first of all, the cut rule is admissible. Generally, this calculus is undecidable, but the fragment without non-local
contraction is PSPACE-complete. These results were obtained by Kanovich, Kuznetsov, Nigam, and Scedrov.

@@

\begin{lemma}
  $I_1 a_1 \cdot I_2 a_2 \leq I (I_1 a_1 \cdot I_2 a_2)$, where $I_i \leq I, i = 1,2$.
\end{lemma}

  $I_1 a_1 \cdot I_2 a_2 \leq I_1 (I_1 a_1) \cdot I_2 (I_2 a_2) \leq I (I_1 a_1) \cdot I (I_2 a_2) \leq I (I_1 a_1 \cdot I_2 a_2)$
\end{proof}

@@

Note that centrality and strong square-increasing yield centrality.

  Let $I$ be an open modality on some unital quantale $\mathcal{Q} = \langle A, \bigvee, \cdot, e \rangle$.
  Then, if $I$ is unital and weak idempotent, then $I$ is central.
\end{lemma}

@@

@@

Hence, $b \cdot I a = I a \cdot b$, so for all $a \in \mathcal{Q}, I a \in \mathcal{Z}(\mathcal{Q})$.

  Let $\mathcal{Q}$ be a quantale and $S \subseteq \mathcal{Q}$ a subquantale,
  then $I : \mathcal{Q} \to \mathcal{Q}$, such that $\Box a = \bigvee \{ s \in S \: | \: s \leq a \}$, is an open modality.
\end{prop}

\begin{proof}
  See Rosenthal.
\end{proof}

  Let $\mathcal{Q}$ be a quantale and $\mathcal{S}_1 \subseteq \mathcal{S}_2 \subseteq \mathcal{Q}$.

  Then $\Box_{\mathcal{S}_1} (a) \leq \Box_{\mathcal{S}_1} (a)$.
\end{prop}

\begin{proof}

Immediatly.
\end{proof}

@@

\begin{enumerate}
  \item $\Box_z (a) = \bigvee \{ s \in S \: | s \leq a, s \in \mathcal{Z}(\mathcal{Q}) \}$;
  \item $\Box_{\mathds{1}} (a) = \bigvee \{ s \in S \: | s \leq a, s \leq \mathds{1} \}$;
  \item $\Box_{idem} (a) = \bigvee \{ s \in S \: | s \leq a, \forall b \in Q, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s\}$;
  \item Operations with combinations of conditions above;
\end{enumerate}
\end{prop}

\begin{proof}
  Immediatly.
\end{proof}

\begin{enumerate}
  \item $\forall a \in \mathcal{Q}, \Box_{\mathds{1}, idem}(a) \leq \Box_z (a)$.
  \item $\forall a \in \mathcal{Q}, \Box_{z, \mathds{1}, idem} = \Box_{\mathds{1}, idem}(a)$
\end{enumerate}

@@

\begin{proof}
  Follows from Lemma 3.
\end{proof}

We define a subexponential interpretation in a quite sophisticated manner. Suppose we have a contravariant map from signature
to the set of subquantales. Here, contravariance denotes that the stronger subexponential index (in sense of preorder) maps
to the weaker subquantale. The second one function maps a subquantale to its quantic conucleus according to the previous
proposition. We match the index pursuant to its subset. In other words, if $s \in W$, then the result of subexponential
interpretation is an unital quantic conucleus, etc. More formally:

@@

Let $\mathcal{Q}$ be a quantale and $\Sigma = \langle \mathcal{I}, \preceq, \mathcal{W}, \mathcal{C}, \mathcal{E} \rangle$ be
a subexponential signature. Suppose, we have $S : \Sigma \to Sub(\mathcal{Q})$, a contravariant map from this subexponential
signature to the set of subquantales of $\mathcal{Q}$, and map $M : \mathcal{P}(\mathcal{Q}) \to \Box_{\mathcal{Q}}$, where
$\Box_{\mathcal{Q}}$ is the set of quantic conuclei on $\mathcal{Q}$. Thus, a subexponential interpretation is the map
$\sigma = M \circ S$ such that:

@@

An interpretation and an entailment relation are defined standardly via valuation map and subexponential interpretation. There occur soundness and completeness theorems:

\begin{theorem}
  $\Gamma \rightarrow A \Rightarrow [\![\Gamma]\!] \leq [\![A]\!]$
\end{theorem}

@@

We consider the promotion case, the rest modal cases are immediatly shown.

@@

By IH, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq [\![A]\!]$. Thus, $\sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!]) \leq \sigma(s)([\![A]\!])$.

By Lemma, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq \sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!])$.

So, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq \sigma(s)([\![A]\!])$.
\end{proof}

@@

  On the other hand, $I(a) \leq a$. So $I(a) \leq \bigvee \{ s \in \mathcal{Q}_{I} \: | \: s \leq a \} = I'(a)$.

@@

Author would like to thank Lev Beklemishev, Stepan Kuznetsov, Fedor Pakhomov, Danyar Shamkanov, and Andre Scedrov
for advice, critique, and comments.

\addcontentsline{toc}{section}{Список использованной литературы}

>>>>

@@

Let us introduce inference polymodal inference rules for non-commutative subexponentials. One can apply substructural
rules only if there is the relevant index on the current modality. Also, the right introduction rule is a sort of the
generalised right $\Box$-introduction a la $K4$. Modality $!_s$ might be introduced on the right only if its index is less than any other subexponential index from the antecedent. That is, if we have already used stronger modality, then we may
apply the weaker one.

@@

@@

Let us consider the current proof-theoretical and algorithmic results on Lambek calculus with additives and subexponentials.
The first of all, the cut rule is admissible. Generally, this calculus is undecidable, but the fragment without non-local
contraction is PSPACE-complete. These results were obtained by Kanovich, Kuznetsov, Nigam, and Scedrov \cite{Sc}.

@@

@@

Here we formulate and prove some techincal lemmas on quantic conuclei that we apply in the proof of soundness and completeness theorems:

\begin{lemma}
  $ $
  Let $\mathcal{Q}$ be a quantale. Let $I$, $I_1$, and $I_2$ be quantic conuclei. Thus, the following properties hold:
  \begin{enumerate}
  \item $I_1 a_1 \cdot I_2 a_2 \leq I (I_1 a_1 \cdot I_2 a_2)$, where $I_i \leq I, i = 1,2$.
  \item If $I$ is unital and weak idempotent, then $I$ is central.
  \item Let $S \subseteq \mathcal{Q}$ be a subquantale, then $I : \mathcal{Q} \to \mathcal{Q}$, such that $\Box a = \bigvee \{ s \in S \: | \: s \leq a \}$, is an open modality.
  \item Let $S_1$ and $S_2$ are subquantales such that $mathcal{S}_1 \subseteq \mathcal{S}_2 \subseteq \mathcal{Q}$. Then $\Box_{\mathcal{S}_1} (a) \leq \Box_{\mathcal{S}_1} (a)$.
  \end{enumerate}
\end{lemma}

@@

\begin{enumerate}
  \item $I_1 a_1 \cdot I_2 a_2 \leq I_1 (I_1 a_1) \cdot I_2 (I_2 a_2) \leq I (I_1 a_1) \cdot I (I_2 a_2) \leq I (I_1 a_1 \cdot I_2 a_2)$
  \item $ $

@@

  Hence, $b \cdot I a = I a \cdot b$, so for all $a \in \mathcal{Q}, I a \in \mathcal{Z}(\mathcal{Q})$.
  \item See Rosental.
  \item Immediately.
\end{enumerate}
\end{proof}

@@

\begin{enumerate}
  \item $I a = \bigvee \{ s \in S \: | \: s \leq a, s \in \mathcal{Z}(\mathcal{Q}) \}$ is a central quantic conucleus
  \item $I a = \bigvee \{ s \in S \: | \: s \leq a, s \leq \mathds{1} \}$ is an unital conucleus
  \item $I a = \bigvee \{ s \in S \: | \: s \leq a, \forall b \in Q, b \cdot s \vee s \cdot b \leq s \cdot b \cdot s\}$ is a generalised square-increasing quantic conucleus
  \item Operations with combinations of conditions above
\end{enumerate}
\end{prop}

\begin{proof}
  Immediately
\end{proof}

\begin{prop}
$ $
Let $\mathcal{Q}$ be a quantale, then:
\begin{enumerate}
  \item Let $I_1$ be a unital and generalised square-increasing and $I_2$ a central, then $\forall a \in \mathcal{Q}, I_1 a \leq I_2 a$
  \item If $I_1$ is GSI, unital, and central, $I_2$ is GSI and unital, then $\forall a \in \mathcal{Q}, I_1 a = I_2 a$
\end{enumerate}

@@

  \begin{enumerate}
    \item Follows from Lemma 1.2.
    \item The right inequation follows from the first of this lemma. The left inequation is obvious.
  \end{enumerate}
\end{proof}

We define a subexponential interpretation in quite sophisticatedly. Suppose we have a contravariant map from a given signature to the set of subquantales. Here, contravariance denotes that the stronger subexponential index (in sense of
preorder) maps to the weaker subquantale. The second one function maps a subquantale to its quantic conucleus according to
the previous proposition. We match the index pursuant to its subset. In other words, if $s \in W$, then the result of s
subexponential interpretation is an unital quantic conucleus, etc. More formally:

@@

Let $\mathcal{Q}$ be a quantale and $\Sigma = \langle \mathcal{I}, \preceq, \mathcal{W}, \mathcal{C}, \mathcal{E} \rangle$
a subexponential signature. Let $S : \Sigma \to Sub(\mathcal{Q})$ be a contravariant map from this subexponential
signature to the set of subquantales of $\mathcal{Q}$ Thus, a subexponential interpretation is the map
$\sigma : \Sigma \to CN(\mathcal{Q})$, where $CN(\mathcal{Q})$ is the set of quantic conuclei on $\mathcal{Q}$, such that:

@@

An interpretation and an entailment relation are defined standardly via valuation map and subexponential interpretation.
The first of all, let us make sure that any subexponential interpretation is sound w.r.t the set of quantic conuclei $CN(\mathcal{Q})$ of the quantale $\mathcal{Q}$.

  Let $\Sigma$ be a subexponential signature and $\sigma : \Sigma \to CN(\mathcal{Q})$ for a contravariant map
  $S : \Sigma \to Sub(\mathcal{Q})$, then:
  \begin{enumerate}
  \item For all $s_1, s_2 \in \Sigma$, if $s_1 \preceq s_2$, then $\sigma(s_2) \leq \sigma(s_1)$.
  In other words, $\sigma$ is a contravariant functor from $\Sigma$ to $Sub(\mathcal{Q})$.
  \item $\sigma(A) = \{ \sigma(s) \: | \: s \in A \}$ is downwardly closed, where $A = \{ \mathcal{W}, \mathcal{C}, \mathcal{E} \}$.
  \item If $s \in \mathcal{W} \cap \mathcal{C}$, then $\sigma(s)$ is a central quantic conucleus.
\end{enumerate}
\end{lemma}

@@

  \begin{enumerate}
  \item Let $s_1, s_2 \in \Sigma$ and $s_1 \leq s_2$. Then $S(s_2) \subseteq S(s_1)$.
  So, $\forall a \in \mathcal{Q}, \bigvee \{ s \in S(s_2) \: | \: s \leq a \} \leq \bigvee \{ s \in S(s_1) \: | \: s \leq a \}$. Thus, $\sigma(s_2) \leq \sigma(s_1)$.
  \item Let $s_1 \in A$ and $s_1 \preceq s_2$. By the previous part, $\sigma(s_2) \leq \sigma(s_1)$. Suppose $\sigma(s_2) \notin \sigma(A)$, then $s_2 \notin A$. But $s_2 \in A$ so far as $A$ is upwardly closed. Contradiction.
  \item Let $s \in \mathcal{W} \cap \mathcal{C}$ and $a \in \mathcal{Q}$,
  then $\sigma(s)(a) = \bigvee \{ x \in S(s) \: | \: x \leq a, x \in \mathcal{Z}(\mathcal{Q}), \text{ and } \forall b \in \mathcal{Q}, (x \cdot b) \vee (b \cdot x) \leq x \cdot b \cdot x \}$. Thus, by Lemma, $\sigma(s)(a) \in \mathcal{Z}(\mathcal{Q})$.

\end{enumerate}
\end{proof}

There occur soundness and completeness theorems.

  Let $\Sigma$ be a subexponential signature, then
  $SMALC_{\Sigma} \vdash \Gamma \rightarrow A \Rightarrow [\![\Gamma]\!] \leq [\![A]\!]$ for some valuation map $f$ and subexponential interpretation $\sigma : \Sigma \to CN(\mathcal{Q})$.
\end{theorem}

By the previous lemma, $\sigma$ is well-defined. Let us consider the case of polymodal promotion.

@@

By IH, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq [\![A]\!]$. Thus,
$\sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!]) \leq \sigma(s)([\![A]\!])$. By Lemma,
$\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq \sigma(s)(\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot
\sigma(s_n) [\![A_n]\!])$. \\
So, $\sigma(s_1)[\![A_1]\!] \cdot \dots \cdot \sigma(s_n) [\![A_n]\!] \leq \sigma(s)([\![A]\!])$.
\end{proof}

@@

  On the other hand, $I(a) \leq a$. So $I(a) \leq \bigvee \{ s \in \mathcal{Q}_{I} \: | \: s \leq a \} = I'(a)$, so far as $I(a) \in \{ s \in \mathcal{Q}_{I} \: | \: s \leq a \}$.

@@

Author would like to thank Lev Beklemishev, Giuseppe Greco, Stepan Kuznetsov, Jean-Francois Mascari, Fedor Pakhomov, Danyar Shamkanov, and Andre Scedrov for advice, critique, and comments.

\addcontentsline{toc}{section}{References}
